Quantitative, Multispecies Monitoring at a Continental Scale

### 2025-02-24 Manuscript_plain_AOS.tex

Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts. For example, \cite{diaz2020a} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material. Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time. We therefore see eDNA as an important datastream that maximizes the information value of existing research cruises, rather than a means of replacing those cruises themselves.

>>>>

Moreover, integrating retrospective eDNA analyses into climate change modeling can enrich predictive frameworks by providing historically grounded baselines of community composition, thereby enhancing our understanding of ecosystem responses to environmental shifts. For example, \cite{diaz2020} demonstrated how archived samples could be leveraged to reconstruct past fish communities using eDNA, thereby complementing traditional monitoring methods and further extending the utility of archived material. Such retrospective applications deepen and extend the value of every research expedition, ensuring that once-collected samples continue to yield valuable ecological insights over time. We therefore see eDNA as an important datastream that maximizes the information value of existing research cruises, rather than a means of replacing those cruises themselves.

### 2025-02-24 Priors table.tex

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
         & \textbf{Description} & \textbf{Prior} \\
&Data & \\
\hline
$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
$Z$ & qPCR amplification (yes=1; no=0) & - \\
$Y$ & qPCR cycle threshold (Ct) & - \\
$R$ & metabarcoding number of reads & - \\
$T$ & total number of reads for all S species & - \\
$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
$V$ & volume filtered and aliquote dilution offset & -\\
$I$ & aliquote dilution factor & -\\

&&\\
&State processes&\\
\hline
$E_{s\neq H}$ & unknown DNA concentration in copies/$\mu$L & $\mathcal{N}(0,5)$\\
$D$ & unknown DNA concentration in copies/L & - \\
$F$ & smoothed DNA concentration in copies/L & - \\
&&\\

&Parameters&\\
\hline
$\psi$& overall mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
$\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
$\sigma_D= (0, +\infty)$& standard deviation of spatial and depth specific random effect & $\mathcal{N}(0,3)$\\

$\delta$& biological replicate random effect & $\mathcal{N}(0,\tau)$\\
$\tau = (0, +\infty)$& standard deviation of biological replicate random effect & $\mathcal{N}(0,0.2)$\\

>>>>

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
         & \textbf{Description} & \textbf{Prior} \\
&Data & \\
\hline
$K$ & known DNA concentration from qPCR standards in ln(copies/$\mu$L) & - \\
$Z$ & qPCR amplification (yes=1; no=0) & - \\
$Y$ & qPCR cycle threshold (Ct) & - \\
$R$ & metabarcoding number of reads & - \\
$T$ & total number of reads for all S species & - \\
$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
$L$ & known DNA concentration from mock commuinties in ln(copies/$\mu$L) & - \\
$V$ & volume filtered and aliquote dilution offset & -\\
$I$ & aliquote dilution factor & -\\

&&\\
&State processes&\\
\hline
$E_{s\neq H}$ & unknown DNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(0,5)$\\
$D$ & unknown DNA concentration in ln(copies/$\mu$L)& - \\
$F$ & smoothed DNA concentration in ln(copies/L) & - \\
&&\\

&Parameters&\\
\hline
$\psi$& overall mean hake eDNA concentration in ln(copies/$\mu$L) & $\mathcal{N}(2,2)$\\
$\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
$\sigma_D $& standard deviation of spatial and depth specific random effect & $\mathcal{TN}(0,3;0, +\infty)$\\

$\delta$& biological replicate random effect & $\mathcal{N}(0,\tau)$\\
$\tau$& standard deviation of biological replicate random effect & $\mathcal{TN}(0,0.2;0, +\infty)$\\
